Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-28 paper/example.tex
Add minor edits to example

From time to time, SW receives an \texttt{AutomationRequest} message,
forwarding it to a route planner (AI). The function of AI is to
compute the flight path (a list of waypoints) for the UAV, outputting
the resulting \emph{mission command} on \texttt{AutomationResponse}.
The waypoint manager (WM) receives the mission command from AI and
starts the UAV flying the mission by putting an \emph{event} on the
\texttt{Start} output port, continuing to issue waypoints to the UAV
flight controller on the \texttt{Waypoint} port as the UAV location
changes.\footnote{\konst{KLS}:What about Alert? When raised, is it
persistent, where is it computed?}  The AI component is third-party
software and WM is a legacy component.

@@

The AGREE contracts for the initial system in \figref{fig:example},
and its implementing components, both assume and guarantee the absence
of any malicious, or unspecified, component behavior.  More precisely,
the contract for SW assumes that its inputs are well-formed and that
there is never more than one automation request pending at a time.
(\emph{Well-formed} generally refers to a syntactic restriction on
data at a port. For example, a waypoint is well-formed if it falls
within bounds for latitude, longitude, and altitude.)  The guarantees
for SW ensure that
\begin{compactitem}
\item a start includes a waypoint;\footnote{\konst{KLS}: needs clarification; Start is just an event port.}
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with air vehicle location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are
well-formed, and they guarantee their outputs are well-formed.  The AI
contract guarantees it only responds to automation requests and always
in the same cycle.  The legacy WM contract guarantees the following if
its input assumptions hold:
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start always comes with a waypoint; and
  \item any further waypoints coincide with air vehicle location updates.
\end{compactitem}
These initial specifications pass AGREE verification, meaning that the
contract composition of the components with the system satisfies all
component input assumptions and system output guarantees.

\subsection{Detecting and addressing vulnerabilities}
A cyber-vulnerability analysis, however, identifies the potential of a
supply chain attack through the AI route planner (provided by a
third-party vendor without source code).  BriefCASE marks the
component as \emph{untrusted} after the analysis, and the contract for
AI is modified by removing all guarantees on its outputs to reflect
its untrusted status.  The AI output is now unconstrained in the
assume-guarantee reasoning of AGREE and is able to generate any value
on its output stream at anytime.

@@

For example, the first violation is that the automation response from
the AI to the WM is no longer guaranteed to be well-formed.  The
consequence of that failing input assumption is that the WM outputs
are no longer guaranteed; they are effectively unconstrained.  These
missing output guarantees from the WM lead to the rest of the failures
in \figref{fig:example-certificate} because the WM provides the system
level outputs.

@@

The system designer now uses BriefCASE to cyber-harden SW by inserting
high-assurance components in the form of a filter and a monitor, as
shown in \figref{fig:hardened}, to isolate the AI component and prevent it from violating the WM component input assumptions.
A filter enforces an invariant over
each datum in the data stream by not forwarding input to its output if
that input violates the filter invariant.  The auto-generated AGREE
specification states that only well-formed inputs are passed to the
output.  The system developer must further define the filtering
policy, but it is usually based on the existing assumptions made by
downstream components that consume the filter output.

A monitor captures a relation on input data over time and is thus able
to reason about the temporal, and other invariant, properties of that input.  A monitor raises
an alert if the specified properties are ever violated.  The
AGREE specification for the monitor in our example states that an
automation response can only be generated in conjunction with an
automation request; and further, that response must come with the
request or in the next step after the request.  As with the filter,
the system designer provides the policy, and that policy is typically
based on existing the assumptions and guarantees of the SW system and its implementing components.

@@

\subsection{Code synthesis}
High-assurance components are automatically synthesized by SPLAT from
the AGREE specifications to equivalent programs in the CakeML
language.
%% Too strong at the moment!
%The synthesis toolchain generates a proof that equates the
%% meaning of the AGREE specification to the behavior of the CakeML
%% program.
 In other words, for any set of input streams that meet the
 component's contract assumptions, the output streams produced from
 the synthesized CakeML code exactly match those defined by the
 high-assurance component's guarantees.  The CakeML compiler rovides
 verified compilation to binaries for several different platforms,
 meaning that the resulting binaries exactly preserve the meaning of
 the original CakeML code \cite{cakeml}.

>>>>

From time to time, SW receives a message on \texttt{AutomationRequest},
forwarding it to a route planner (AI). The function of AI is to 
compute the flight path (a list of waypoints) for the UAV, outputting
the resulting \emph{mission command} on \texttt{AutomationResponse}.
The waypoint manager (WM) receives the mission command from AI and
starts the UAV flying the mission by putting an \emph{event} on the
\texttt{Start} output port, continuing to issue waypoints to the UAV
flight controller on the \texttt{Waypoint} port as the UAV location
updates with messages on \texttt{AirVehicleLocation}.
An event is repeatedly generated on \texttt{Alert} if there is no
mission command from a request. The AI component is third-party
software and WM is a legacy component.

@@

The AGREE contracts for the initial system in \figref{fig:example},
and its implementing components, both assume and guarantee the absence
of any malicious, or unspecified, component behavior.  More precisely,
the contract for SW assumes that its inputs are \emph{well-formed} and that
there is never more than one automation request pending at a time.
Well-formed generally refers to a syntactic restriction on
data at a port. For example, a waypoint is well-formed if it falls
within bounds for latitude, longitude, and altitude.)  The guarantees
for SW ensure that
\begin{compactitem}
\item a start coincides with a new waypoint message being output;
\item a start is within one cycle of an automation request and if not, then it persistently alerts;
\item new waypoints coincide with air vehicle location updates; and
\item all outputs are well-formed.
\end{compactitem}

The contracts for the sub-components assume their inputs are
well-formed, and they guarantee their outputs are well-formed.  The AI
contract guarantees it only responds to automation requests and always
in the same cycle.  The legacy WM contract guarantees the following if
its input assumptions hold:
\begin{compactitem}
  \item it generates a start from a response from the AI and always in the same cycle;
  \item the start always coincides with a waypoint being output; and
  \item any further waypoints coincide with air vehicle location updates.
\end{compactitem}
These initial specifications pass AGREE verification, meaning that the
contract composition of the components with the system satisfies all
component input assumptions and system output guarantees.

\subsection{Detecting and addressing vulnerabilities}
A cyber-vulnerability analysis, however, identifies the potential of a
supply chain attack through the AI route planner (provided by a
third-party vendor without source code).  BriefCASE marks the
component as \emph{untrusted} after the analysis, and the contract for
AI is modified by removing all guarantees on its outputs to reflect
its untrusted status.  The AI output is now unconstrained in the
assume-guarantee reasoning of AGREE and is able to generate any value
on its output at anytime.

@@

For example, the first violation is that the mission command on \texttt{AutomationResponse} from
the AI to the WM is no longer guaranteed to be well-formed.  The
consequence of that failing input assumption is that the WM outputs
are no longer guaranteed; they are effectively unconstrained.  These
missing output guarantees from the WM lead to the rest of the failures
in \figref{fig:example-certificate} because the WM provides the system
level outputs.

@@

The system designer now uses BriefCASE to cyber-harden SW by inserting
high-assurance components in the form of a filter and a monitor, as
shown in \figref{fig:hardened}, to isolate the AI component and prevent it from violating the WM component input assumptions.
A filter enforces an invariant over
each datum in the data stream by not forwarding input to its output if
that input violates the filter invariant.  The auto-generated AGREE
code-contract specification states that only well-formed inputs are passed to the
output.  The system developer must further define the filtering
policy, but it is usually based on the existing assumptions made by
downstream components that consume the filter output.

A monitor captures a relation on input data over time and is thus able
to reason about the temporal, and other invariant, properties of that input.  A monitor raises
an alert if the specified properties are ever violated.  The
AGREE code-contract specification for the monitor in our example states that an
automation response can only be generated in conjunction with an
automation request; and further, that response must come with the
request or in the next step after the request.  As with the filter,
the system designer provides the policy, and that policy is typically
based on existing the assumptions and guarantees of the SW system and its implementing components.
AGREE code-contract specifications are discussed in detail in \secref{sec:code-contracts}.

@@

\subsection{Code synthesis}
High-assurance components are automatically synthesized by SPLAT from
the AGREE code-contract specifications to equivalent programs in the CakeML
language.
%% Too strong at the moment!
%The synthesis toolchain generates a proof that equates the
%% meaning of the AGREE specification to the behavior of the CakeML
%% program.
 In other words, for any set of input streams that meet the
 component's contract assumptions, the output streams produced from
 the synthesized CakeML code exactly match those defined by the
 high-assurance component's guarantees.  The CakeML compiler provides
 verified compilation to binaries for several different platforms,
 meaning that the resulting binaries exactly preserve the meaning of
 the original CakeML code \cite{cakeml}.
